---
title: "Pb_Spz_10X_4"
author: "Anthony Rubertoand Caitlin Bourke"
output: html_document
---

  i. Load libraries.

```{r load libraries}

suppressPackageStartupMessages({
  library(SingleCellExperiment)
  library(dplyr)
  library(scater)
  library(pipeComp)
  library(scDblFinder)
  library(ggpubr)
  library(RColorBrewer)
  library(Seurat)
  })

```
  ii. Load color palette.
  
```{r load color palette}

palette<-c( '#3cb44b', "#e6194b", '#4363d8','#ffe119','#f58231', '#911eb4', '#46f0f0', '#800000', '#ffd8b1', '#fabebe', '#008080', '#e6beff', '#9a6324', 'grey60', '#f032e6', '#aaffc3', '#808000', '#bcf60c', '#000075', '#808080', '#ffffff', '#000000')

```

The files used in this document were downloaded from MCA's dedicated Github page [1]. 

1. Load MCA SmartSeq-2 data in to R, and subset to only keep pre-vertebrate life-cycle stages of the parasite.

```{r load MCA PbANKA SS2 data}

molecules <- read.table("~/Documents/Single_cell_practice/MalariaAtlas/SS2_counts.txt", header = TRUE, sep = ",", row.names=1, stringsAsFactors = TRUE)
anno <- read.delim("~/Documents/Single_cell_practice/Git_10X_Pb_Spz/SS2_pheno.txt", header = TRUE, sep = ",")
cols <- c("bbSpz" = "navy", "EEF"="darkorange", "Merozoite"="lightpink", "oocyst"="steelblue", "ook" = "turquoise4", "Ring"="hotpink", "sgSpz"= "royalblue", "Schizont" = "violetred", "Male"="purple", "Female"="purple4", "ookoo" = "mediumturquoise", "Trophozoite"="violet")

mca.qc.ss <- SingleCellExperiment(assays = list(
  counts = as.matrix(molecules)), 
  colData = anno,
  rowData = data.frame(gene_names = rownames(molecules))
  )

mca.qc.ss <- mca.qc.ss[, (colData(mca.qc.ss)$ShortenedLifeStage3 == "bbSpz") | 
                       (colData(mca.qc.ss)$ShortenedLifeStage3 == "sgSpz") |
                       (colData(mca.qc.ss)$ShortenedLifeStage3 == "Oocyst") |
                       (colData(mca.qc.ss)$ShortenedLifeStage3 == "Ookinete")]

```

2. Prep matrix in a similar manner used in 'Pb_Spz_10X.rmd'.

```{r prep matrices MCA PbANKA SS2}

Pb_v46_genes<- read.delim("/Users/aruberto/Documents/Projects/Pb_spz_10X_sc/Analyses/R/PbANKA_v46_info.txt") %>% distinct()

rowData(mca.qc.ss)$Description <- Pb_v46_genes$Description[match(rownames(mca.qc.ss), Pb_v46_genes$Gene)]
rowData(mca.qc.ss)$GeneDescription <- Pb_v46_genes$GeneDescription[match(rownames(mca.qc.ss), Pb_v46_genes$Gene)]
rowData(mca.qc.ss)$ID<-rownames(mca.qc.ss)
colData(mca.qc.ss)$Replicate<-"Biological"
colData(mca.qc.ss)$Sample<-"SS2"
colData(mca.qc.ss)$sample_id<-paste(colData(mca.qc.ss)$sample_id,"_",colData(mca.qc.ss)$Sample, sep = "")
colnames(mca.qc.ss)<-colData(mca.qc.ss)$sample_id
colData(mca.qc.ss)$X <- NULL
colData(mca.qc.ss)

```

3. Find doublets.

```{r find doublets in MCA Pb ANKA SS2 dataset}

set.seed(6969)

mca.qc.ss <- scDblFinder(mca.qc.ss, verbose=T, score= "hybrid")

colData(mca.qc.ss)

```

4. Add per cell metrics.

  i. Identify rRNA genes

```{r rRNA for MCA Pb SS2}

rRNA_all<-c("PBANKA_1024351","PBANKA_1024331","PBANKA_0700721","PBANKA_MIT00300","PBANKA_MIT00700","PBANKA_MIT02500",
            "PBANKA_MIT03100","PBANKA_MIT02400","PBANKA_MIT02200","PBANKA_MIT00800","PBANKA_MIT01200","PBANKA_MIT01700",
            "PBANKA_MIT01000","PBANKA_MIT00900","PBANKA_0700761","PBANKA_0622961","PBANKA_MIT01400","PBANKA_API00051",
            "PBANKA_1245821","PBANKA_MIT01500","PBANKA_MIT00100","PBANKA_MIT02300","PBANKA_MIT01300","PBANKA_MIT03600",
            "PBANKA_MIT03200","PBANKA_MIT00600","PBANKA_MIT02600","PBANKA_MIT02900","PBANKA_MIT00200","PBANKA_0622941",
            "PBANKA_MIT03000","PBANKA_API00095","PBANKA_MIT02000","PBANKA_0622921","PBANKA_MIT02800","PBANKA_0521241",
            "PBANKA_1024341","PBANKA_MIT00400","PBANKA_0521221","PBANKA_MIT03500","PBANKA_0521261","PBANKA_MIT03700",
            "PBANKA_MIT01600","PBANKA_1245841","PBANKA_1245861","PBANKA_MIT03300","PBANKA_MIT03400","PBANKA_API00055",
            "PBANKA_API00011","PBANKA_MIT00500","PBANKA_MIT02700","PBANKA_MIT02100","PBANKA_0700741")

```

  ii. Add rRNA and mito information to SCE object.

```{r add mito and rRNA to MCA Pb SS2}

# Get Mitochondrial genes and rRNA for QC:

mca.mt_genes <- grepl("^PBANKA_MIT", rowData(mca.qc.ss)$ID)
mca.ribo_genes <-rRNA_all[(rRNA_all %in% rowData(mca.qc.ss)$ID)]
mca.feature_ctrls <- list(mito = rowData(mca.qc.ss)$ID[mca.mt_genes], ribo = mca.ribo_genes)

lapply(mca.feature_ctrls, head)

```
  iii. Calculate per cell QC metrics.

```{r QC per cell metrics}

mca.qc.ss <- addPerCellQC(mca.qc.ss, subsets = mca.feature_ctrls)
mca.qc.ss <- addPerFeatureQC(mca.qc.ss, detection_limit = 0)

```

5. Filter potential doublets.

```{r filter high count MCA Pb SS2}

mca.qc.ss$is_outlierCounts <- isOutlier(metric = mca.qc.ss$sum, nmads = 3, type = "higher", log = TRUE)

mca.qc.ss <- mca.qc.ss[, !mca.qc.ss$is_outlierCounts]

dim(mca.qc.ss)

```

6. Filter low detected genes.

```{r filter low detect genes MCA Pb SS2}

dim(mca.qc.ss)

mca.qc.ss<-mca.qc.ss[rowSums(counts(mca.qc.ss) > 2) > 1, ]

dim(mca.qc.ss)

```

7. Recalculate per-cell metrics post-processing.

  i. Basics.

```{r reculaculate QC per cell metrics for MCA Pb SS2}

mca.mt_genes <- grepl("^PBANKA_MIT", rowData(mca.qc.ss)$ID)
mca.ribo_genes <-rRNA_all[(rRNA_all %in% rowData(mca.qc.ss)$ID)]
mca.feature_ctrls <- list(mito = rowData(mca.qc.ss)$ID[mca.mt_genes],
                      ribo = mca.ribo_genes)

lapply(mca.feature_ctrls, head)

colData(mca.qc.ss)[,c("preFilterSum", "preFilterDetected")]<-colData(mca.qc.ss)[,54:55]
colData(mca.qc.ss)[,c(37:44, 54:65)]<-NULL
colData(mca.qc.ss)

rowData(mca.qc.ss)[,c("preFilterMean", "preFilterDetected")]<-rowData(mca.qc.ss)[,5:6]
rowData(mca.qc.ss)[,5:6]<-NULL
rowData(mca.qc.ss)

mca.qc.ss <- addPerCellQC(mca.qc.ss, subsets = mca.feature_ctrls)
colData(mca.qc.ss)

# Note: all both scDblFinder and is_OutlierCounts metrics carried forward are based on the PRE-FILTERED cell numbers.

mca.qc.ss <- addPerFeatureQC(mca.qc.ss, detection_limit = 0)

```

  ii. PipeComp supplemental.
  
```{r add supplementary column metadata to MCA Pb SS2}

source(system.file("extdata", "scrna_alternatives.R", package="pipeComp"))

add_supp_meta <- function(ds){
  ds$total_features <- ds$detected
  ds$log10_total_features <- log10(ds$detected)
  ds$total_counts <- ds$sum
  ds$log10_total_counts <- log10(ds$sum+1)
  ds$featcount_ratio <- ds$log10_total_counts/ds$log10_total_features
  ds$featcount_dist <- getFeatCountDist(ds)
  ds$pct_counts_top_50_features <- ds$percent_top_50
  ds
}

mca.qc.ss<-add_supp_meta(mca.qc.ss)

colData(mca.qc.ss)

```


8. SCE to Seurat for data reduction and clustering.

 i. ensure compatibility I 

```{r sce to Seurat MCA Pb SS2}

mca.mq.stage<-as.Seurat(mca.qc.ss, data = NULL)

mca.mq.stage$nCount_RNA <-mca.mq.stage$sum

mca.mq.stage$nFeature_RNA <-mca.mq.stage$detected

```

  ii. ensure compatibility II
  
```{r rRNA and UIS calling}

# vectors containing rRNA and UIS genes 

rRNA_all<-c("PBANKA-1024351","PBANKA-1024331","PBANKA-0700721","PBANKA-MIT00300","PBANKA-MIT00700","PBANKA-MIT02500",
            "PBANKA-MIT03100","PBANKA-MIT02400","PBANKA-MIT02200","PBANKA-MIT00800","PBANKA-MIT01200","PBANKA-MIT01700",
            "PBANKA-MIT01000","PBANKA-MIT00900","PBANKA-0700761","PBANKA-0622961","PBANKA-MIT01400","PBANKA-API00051",
            "PBANKA-1245821","PBANKA-MIT01500","PBANKA-MIT00100","PBANKA-MIT02300","PBANKA-MIT01300","PBANKA-MIT03600",
            "PBANKA-MIT03200","PBANKA-MIT00600","PBANKA-MIT02600","PBANKA-MIT02900","PBANKA-MIT00200","PBANKA-0622941",
            "PBANKA-MIT03000","PBANKA-API00095","PBANKA-MIT02000","PBANKA-0622921","PBANKA-MIT02800","PBANKA-0521241",
            "PBANKA-1024341","PBANKA-MIT00400","PBANKA-0521221","PBANKA-MIT03500","PBANKA-0521261","PBANKA-MIT03700",
            "PBANKA-MIT01600","PBANKA-1245841","PBANKA-1245861","PBANKA-MIT03300","PBANKA-MIT03400","PBANKA-API00055",
            "PBANKA-API00011","PBANKA-MIT00500","PBANKA-MIT02700","PBANKA-MIT02100","PBANKA-0700741")

UIS_genes<-c("PBANKA-0205800","PBANKA-1328000","PBANKA-1400800","PBANKA-0501200","PBANKA-0211300","PBANKA-1454900","PBANKA-1233500","PBANKA-1128100",
             "PBANKA-1206300","PBANKA-0506200","PBANKA-1003400","PBANKA-0511400","PBANKA-1232400","PBANKA-1346300","PBANKA-1321400","PBANKA-1311000",
             "PBANKA-0602000","PBANKA-1413400","PBANKA-0914400","PBANKA-1320100",
             "PBANKA-1113800","PBANKA-1435900","PBANKA-1017500","PBANKA-1020200")

# percentage of reads containing encoding for rRNA per cell

mca.mq.stage.genes<-mca.mq.stage@assays$RNA@counts@Dimnames[[1]]
rRNA_all %in% mca.mq.stage.genes
mca.mq.stage.rRNA<-rRNA_all[(rRNA_all %in% mca.mq.stage.genes)]
mca.mq.stage[["percent.rRNA"]]<-PercentageFeatureSet(mca.mq.stage, features = mca.mq.stage.rRNA)

# percentage of reads containing encoding for UIS per cell

UIS_genes %in% mca.mq.stage.genes
mca.mq.stage.UIS<-UIS_genes[(UIS_genes %in% mca.mq.stage.genes)]
mca.mq.stage$percentUIS<-PercentageFeatureSet(object = mca.mq.stage, features = mca.mq.stage.UIS, assay = 'RNA')

```

9. Save MCA Pb ANKA mosquito development data.

This Seurat object was used for the generation of the UMAPs of the integrated 10X and MCA data in Figure 3.

```{r save MCA pre-vertebrate host Seurat object}

saveRDS(mca.mq.stage, "output/MS/4_MCA_SS2_seurat.rds")

```

Links:

[1] https://github.com/vhowick/MalariaCellAtlas/tree/master/Expression_Matrices/Smartseq2


