---
title: "Pb_Spz_10X_6"
author: "Anthony Ruberto and Caitlin Bourke"
output: html_document
---

  i. Load libraries.

```{r load libraries}

suppressPackageStartupMessages({
  library(SingleCellExperiment)
  library(Seurat)
  library(ggpubr)
  library(slingshot)
  library(tradeSeq)
  })

```

  ii. Load color palettes
  
```{r color palette}

palette<-c( '#3cb44b', "#e6194b", '#4363d8','#f58231','#911eb4', 'lightgrey','#ffe119', '#46f0f0', '#800000', '#ffd8b1', '#fabebe', '#008080', '#e6beff', '#9a6324', 'grey60', '#f032e6', '#aaffc3', '#808000', '#bcf60c', '#000075', '#808080', '#ffffff', '#000000')

palette2<-c('#f58231', '#911eb4', '#46f0f0', '#800000', '#008080','grey60', '#f032e6','#808000','red', '#aaffc3',  '#bcf60c', '#000075', '#808080', '#ffffff', '#000000','#3cb44b', "#e6194b", '#4363d8','#ffe119','#ffd8b1', '#fabebe' ,'#e6beff', '#9a6324')

```

1. Upload 10X sporozoite data from Pb_Spz_10X_5.rmd.

```{r upload 10X data}

Pb.integrated.10X.only<-readRDS("output/MS/5_Pb_integrated_10X_only.rds")

```

2. Reformat data so that it is compatible with Slingshot.

```{r Seurat to Slingshot}

dimred <- Pb.integrated.10X.only@reductions$umap@cell.embeddings
clustering <- Pb.integrated.10X.only$integrated_snn_res.0.7
counts <- as.matrix(Pb.integrated.10X.only@assays$RNA@counts[Pb.integrated.10X.only@assays$integrated@var.features, ])

```

3. Determine the lineages in the data. 

```{r lineage identification}

# Run default Slingshot lineage identification

set.seed(1)
lineages <- getLineages(data = dimred, clusterLabels = clustering)
lineages

```

4. Find principal curves.

```{r find principal curves}

curvesALL <- getCurves(lineages, approx_points = F, allow.breaks = FALSE)
curvesALL

```

5. Plot principal curves on UMAP.

```{r plot Principal curves}

# Figure 4a.

plot(dimred[,c(1,2)], col = palette2[clustering], asp = 1, pch = 16, cex = .5)
lines(curvesALL, lwd = 7, col = c("black","green", "blue", 'red'))

```

6. Fit GAM - takes some time. Grab a coffee.

```{r GAM fitting SG 10X scRNA-seq}

sceALL10Xspz <- fitGAM(counts = counts, sds = curvesALL)

```

Once the GAM fitting is complete, use tradeSEQ [1]  on the sceALL10Xspz object to:

1) Assess differential expression pattern between lineages - patternTest().
2) Perform statistical test to check for DE between final stages of every lineage - diffEndTest().

7. Save RDS.

```{r save SCE pseudotime object}

saveRDS(sceALL10Xspz, "output/MS/6_Pb_integrated_10X_only_pseudotime.rds")

```

